---
title: "Covid19_data"
author: "TC"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sessionInfo()
library(tidyverse)
library(lubridate)
```

## Setup
Make sure to library in `library(tidyverse)` in the console.
[Data from this link.](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)

```{r import_data, message = FALSE}
url_in <- "https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/"

file_names <- c("time_series_covid19_confirmed_global.csv", "time_series_covid19_deaths_global.csv", "time_series_covid19_confirmed_US.csv", "time_series_covid19_deaths_US.csv", "time_series_covid19_recovered_global.csv")

urls <- str_c(url_in, file_names)


global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
global_recovered <- read_csv(urls[5])
```

# Tidy up data
Tidying and reformatting are needed. But in particular, we will be focusing on reported cases, reported deaths, and reported recovered. Most notably, in the recovered data, all data seems to have stopped being recorded on the date of 2021-08-05 for all countries. This could be due to a lack of data, but also further research tells us that the first vaccine was released around this time. So there is a chance that the researchers stopped recording recovered cases due to this.

```{r tidy_data, message = TRUE}
global_deaths <- global_deaths %>%
  pivot_longer(cols = -c('Province/State',
                          'Country/Region'),
              names_to = "date",
               values_to = "deaths")
 global_cases <- global_cases %>%
   pivot_longer(cols = -c('Province/State',
                           'Country/Region', Lat, Long),
               names_to = "date",
                values_to = "cases") %>%
   select(-c(Lat,Long))
 
 global_recovered <- global_recovered %>%
   pivot_longer(cols = -c('Province/State',
                           'Country/Region', Lat, Long),
               names_to = "date",
                values_to = "cases") %>%
   mutate(date = mdy(date)) %>%
   rename(recovered = 'cases') %>%
   rename(Country_Region = 'Country/Region') %>%
   rename(Province_State = 'Province/State') %>%
   select(-c(Lat,Long)) %>%
   filter(date < "2021-08-05")
 
 global <- global_cases %>%
   full_join(global_deaths) %>%
   rename(Country_Region = 'Country/Region',
          Province_State = 'Province/State') %>%
   mutate(date = mdy(date))
 global <- global %>% filter(cases > 0)
 
 global <- global %>%
     full_join(global_recovered)
 
 global <- global %>%
    unite("Combined_Key",
          c(Province_State, Country_Region),
          sep = ", ",
          na.rm = TRUE,
          remove = FALSE)

uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url) %>%
  select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

global <- global %>%
  left_join(uid, by = c("Province_State", "Country_Region")) %>%
  select(-c(UID, FIPS)) %>%
  select(Province_State, Country_Region, date, 
         cases, deaths, recovered, Population, Combined_Key)

```

# Visualizing data
Looking at this data globally is too much to absorb all at once (see All cases worldwide), so we shall focus on the United Kingdom. Plotted are the prognosis as a percentage of reported cases. One would assume that the total recovered and total deaths would add up to total cases, but it is evident that the sum falls short. This is probably a result of incomplete data, subjects not following up with their prognosis, or could even be due to entities not faithfully reporting all deaths and recovered cases.

```{r visualize_data}
global_total <- global %>%
  filter(date < "2021-08-05") %>%
  group_by(Country_Region) %>%
  summarize(deaths = max(deaths, na.rm = TRUE), cases = max(cases, na.rm = TRUE), recovered = max(recovered, na.rm = TRUE),
            population = max(Population),
            cases_per_thou = 1000*cases/population,
            recovered_per_thou = 1000*recovered/population,
            deaths_per_thou = 1000*deaths/population) %>%
  filter(cases > 0, population > 0)

global_total <- global_total %>%
  mutate(cases_by_cases = cases/cases) %>%
  mutate(deaths_by_cases = deaths/cases) %>%
  mutate(recovered_by_cases = recovered/cases)

global_total_long <- global_total %>%
  select('Country_Region', 'cases_by_cases', 'recovered_by_cases', 'deaths_by_cases') %>%
  pivot_longer(cols = c('cases_by_cases', 'recovered_by_cases', 'deaths_by_cases'),
                  names_to = "category", values_to = "percentage")
ggplot(global_total_long, aes(x=Country_Region, y=percentage, fill=category)) + 
  geom_bar(stat = "identity", position = "dodge") + theme(axis.text.x = element_text(angle=90), legend.position = "top") +
  labs(title = str_c("All cases worldwide"), y=NULL)

country = "United Kingdom"
global %>%
    filter(Country_Region == country) %>%
    filter(cases > 0) %>%
    ggplot(aes(x = date, y = cases)) + 
    geom_line(aes(color = "cases")) + 
    geom_point(aes(color = "cases")) + 
    geom_line(aes(y=deaths, color="deaths")) + 
    geom_point(aes(y=deaths, color = "deaths")) + 
    geom_line(aes(y=recovered, color="recovered")) + 
    geom_point(aes(y=recovered, color = "recovered")) + 
    scale_y_log10() + 
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 90)) + 
    labs(title = str_c("COVID19 in ", country), y=NULL)

UK_total <- global %>%
  filter(Country_Region == country) %>%
  group_by(Province_State) %>%
  summarize(deaths = max(deaths, na.rm = TRUE), cases = max(cases, na.rm = TRUE), recovered = max(recovered, na.rm = TRUE),
            population = max(Population),
            cases_per_thou = 1000*cases/population,
            recovered_per_thou = 1000*recovered/population,
            deaths_per_thou = 1000*deaths/population) %>%
  filter(cases > 0, population > 0) %>%
  mutate(cases_by_cases = cases/cases) %>%
  mutate(deaths_by_cases = deaths/cases) %>%
  mutate(recovered_by_cases = recovered/cases)

UK_total_long <- UK_total %>%
  select('Province_State', 'cases_by_cases', 'recovered_by_cases', 'deaths_by_cases') %>%
  pivot_longer(cols = c('cases_by_cases', 'recovered_by_cases', 'deaths_by_cases'),
                  names_to = "category", values_to = "percentage")
ggplot(UK_total_long, aes(x=Province_State, y=percentage, fill=category)) + 
  geom_bar(stat = "identity", position = "dodge") + theme(axis.text.x = element_text(angle=90), legend.position = "top") + 
  labs(title = str_c("COVID19 in ", country, " as percentage"), y=NULL)
```

## Modeling data
Here we try to see if we can predict rate of recovery from global data. We test to see if a linear relationship can be achieved between reported cases and recovered cases. Then, we create a model and plot its results alongside actual data to see its accuracy. Interestingly, we see that the model (red) is close to the actual data, but does deviate as more cases were reported.

```{r modeling_data}
mod_global <- lm(recovered_per_thou ~ cases_per_thou, data = global_total)
global_total %>% slice_min(cases_per_thou)
global_total %>% mutate(pred = predict(mod_global))

global_w_pred <- global_total %>% mutate(pred = predict(mod_global))
global_w_pred %>% ggplot() +
  geom_point(aes(x=cases_per_thou, y = recovered_per_thou), color = "blue") + 
  geom_point(aes(x=cases_per_thou, y=pred), color = "red") +
  theme(legend.position = "bottom") + 
  labs(color = "Legend") +
  labs(title = str_c("Prediction of recovered cases globally"), y=NULL)
```